---
title: "mini-project-2"
author: "Victor Yuan"
date: "22/09/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Version 1.1**

*To complete this milestone, edit this `.rmd` file directly. Fill in the sections that are commented out with `<!--- start your work here--->`. When you are done, make sure to knit to an `.md` file by changing the YAML settings, before submitting a tagged release on canvas.*

# Welcome back to your mini data analysis project!

This time, we will explore more in depth the concept of *tidy data*, and hopefully investigate further into your research questions that you defined in milestone 1.

Begin by loading your data and the tidyverse package below: 

```{r, message = FALSE}
library(datateachr) # <- contains the data you picked!
library(tidyverse)
```

# Learning Objectives

By the end of this milestone, you should:

+ Become familiar with manipulating and summarizing your data in tibbles using `dplyr`, with a research question in mind
+ Understand what *tidy* data is, and how to create it. In milestone 3, we will explore when this might be useful
+ Generate a reproducible and clear report using R Markdown
+ Gain a greater understanding of how to use R to answer research questions about your data

**Things to keep in mind**

+ Remember to document your code, be explicit about what you are doing, and write notes in this markdown document when you feel that context is required. Create your analysis as if someone else will be reading it! **There will be 2.5 points reserved for reproducibility, readability, and repo organization.**

+ Before working on each task, you should always keep in mind the specific **research question** that you're trying to answer.

# Task 1: Process and summarize your data (15 points)

From milestone 1, you should have an idea of the basic structure of your dataset (e.g. number of rows and columns, class types, etc.). Here, we will start investigating your data more in-depth using various `dplyr` functions. 

### 1.1 (2.5 points) 

First, write out the 4 research questions you defined in milestone 1 were. This will guide your work through milestone 2:

<!-------------------------- Start your work below ---------------------------->
1) Do the variables radius_mean and radius_worst have strong corrolation? What is their relationship?
2) Are there variables that have relationship with diagnosis?
3) Do the variables concavity_mean and compactness_mean have strong corrolation? What is their relationship?
4) Can we predict diagnosis by a variable or a group of variables?
<!----------------------------------------------------------------------------->

### 1.2 (10 points)

Now, for each of your four research questions, choose one task from options 1-4 (summarizing), and one other task from 4-8 (graphing). You should have 2 tasks done for each research question (8 total). Make sure it makes sense to do them! (e.g. don't use a numerical variables for a task that needs a categorical variable.). Comment on why each task helps (or doesn't!) answer the corresponding research question. 

Ensure that the output of each operation is printed! 



**Summarizing:**

1. Compute the *range*, *mean*, and *two other summary statistics* of **one numerical variable** across the groups of **one categorical variable** from your data.
2. Compute the number of observations for at least one of your categorical variables. Do not use the function `table()`!
3. Create a categorical variable with 3 or more groups from an existing numerical variable. You can use this new variable in the other tasks! *An example: age in years into "child, teen, adult, senior".*
4. Based on two categorical variables, calculate two summary statistics of your choosing.

**Graphing:**

5. Create a graph out of summarized variables that has at least two geom layers.
6. Create a graph of your choosing, make one of the axes logarithmic, and format the axes labels so that they are "pretty" or easier to read. 
7. Make a graph where it makes sense to customize the alpha transparency.
8. Create 3 histograms out of summarized variables, with each histogram having different sized bins. Pick the "best" one and explain why it is the best.

Make sure it's clear what research question you are doing each operation for!

<!------------------------- Start your work below ----------------------------->
```{r}
cancer_sample[1:20,]
```

1) Do the variables radius_mean and radius_worst have strong corrolation? What is their relationship?

```{r}
cancer <- cancer_sample 

range_radius_mean=max(cancer$texture_mean)-min(cancer$texture_mean)

mean_radius_mean=mean(cancer$texture_mean)

median_radius_mean=median(cancer$texture_mean)

sd_radius_mean=sd(cancer$texture_mean)

cat("The range of radius_mean is equal to ", range_radius_mean, "\n")
cat("The mean of radius_mean is equal to ", mean_radius_mean, "\n") 
cat("The median of radius_mean is equal to ", median_radius_mean, "\n") 
cat("The standard deviation of radius_mean is equal to ", sd_radius_mean, "\n") 

```

```{r}
ggplot(data = cancer, mapping = aes(x = radius_mean, y = radius_worst)) + 
  geom_point() + 
  geom_smooth()
```


2) Are there variables that have relationship with diagnosis?

```{r}
ggplot(data = cancer, mapping = aes(x = radius_mean, y = radius_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```


```{r}
ggplot(data = cancer, mapping = aes(x = symmetry_mean, y = symmetry_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

```{r}
ggplot(data = cancer, mapping = aes(x = texture_mean, y = texture_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

```{r}
ggplot(data = cancer, mapping = aes(x = perimeter_mean, y = perimeter_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

```{r}
ggplot(data = cancer, mapping = aes(x = area_mean, y = area_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

```{r}
ggplot(data = cancer, mapping = aes(x = smoothness_mean, y = smoothness_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

```{r}
ggplot(data = cancer, mapping = aes(x = compactness_mean, y = compactness_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

```{r}
ggplot(data = cancer, mapping = aes(x = concavity_mean, y = concavity_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

```{r}
ggplot(data = cancer, mapping = aes(x = concave_points_mean, y = concave_points_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

```{r}
ggplot(data = cancer, mapping = aes(x = symmetry_mean, y = symmetry_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

```{r}
ggplot(data = cancer, mapping = aes(x = fractal_dimension_mean, y = fractal_dimension_worst)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

3) Do the variables concavity_mean and compactness_mean have strong corrolation? What is their relationship?

```{r}
cancer <- cancer_sample 

range_concavity_mean=max(cancer$concavity_mean)-min(cancer$concavity_mean)

mean_concavity_mean=mean(cancer$concavity_mean)

median_concavity_mean=median(cancer$concavity_mean)

sd_concavity_mean=sd(cancer$concavity_mean)

cat("The range of concavity_mean is equal to ", range_concavity_mean, "\n")
cat("The mean of concavity_mean is equal to ", mean_concavity_mean, "\n") 
cat("The median of concavity_mean is equal to ", median_concavity_mean, "\n") 
cat("The standard deviation of concavity_mean is equal to ", sd_concavity_mean, "\n") 

```

```{r}

```


```{r}
ggplot(data = cancer, mapping = aes(x = concavity_mean, y = compactness_mean)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

4. Based on the graphs, we can predict diagnosis by some variables like radius_mean, area_mean,.... It seems that we only need to use one of the variables radius_mean, area_mean, perimeter_mean. So let's look at the following graph 


```{r}
ggplot(data = cancer, mapping = aes(x = radius_mean, y = concave_points_mean)) + 
  geom_point(mapping = aes(color = diagnosis)) + 
  geom_smooth()
```

It seems we can use the variables radius_mean and concave_points_mean to predict diagnosis.


<!----------------------------------------------------------------------------->

### 1.3 (2.5 points)

Based on the operations that you've completed, how much closer are you to answering your research questions? Think about what aspects of your research questions remain unclear. Can your research questions be refined, now that you've investigated your data a bit more? Which research questions are yielding interesting results?

<!------------------------- Write your answer here ---------------------------->

I find out that many variables have strong correlation with each other. So, in order to predict the diagnosis, we only need few variables and can ignore other variables who have strong correlation with them. In particular, few variabels such as radius_mean, concave_points_mean and concavity_mean indicate better the type of diagnosis.

<!----------------------------------------------------------------------------->

# Task 2: Tidy your data (12.5 points)

In this task, we will do several exercises to reshape our data. The goal here is to understand how to do this reshaping with the `tidyr` package.

A reminder of the definition of *tidy* data:

- Each row is an **observation**
- Each column is a **variable**
- Each cell is a **value**

*Tidy'ing* data is sometimes necessary because it can simplify computation. Other times it can be nice to organize data so that it can be easier to understand when read manually. 

### 2.1 (2.5 points)

Based on the definition above, can you identify if your data is tidy or untidy? Go through all your columns, or if you have >8 variables, just pick 8, and explain whether the data is untidy or tidy.

<!--------------------------- Start your work below --------------------------->
The data is tidy because
- Each variable has its own column.
- Each observation has its own row.
- Each value has its own cell.

<!----------------------------------------------------------------------------->

### 2.2 (5 points)

Now, if your data is tidy, untidy it! Then, tidy it back to it's original state.

If your data is untidy, then tidy it! Then, untidy it back to it's original state.

Be sure to explain your reasoning for this task. Show us the "before" and "after".

<!--------------------------- Start your work below --------------------------->
```{r}
(cancer_sample_untidy <- cancer_sample %>%
    pivot_wider(id_cols = c(-radius_mean), 
                names_from = diagnosis,
                values_from = radius_mean)
 )
```

This dataframe is untidy because there is no value in many cells for columns M and B.


```{r}
(cancer_sample_tidy <- cancer_sample_untidy %>% 
  pivot_longer(cols = c(M,B), 
               values_drop_na = TRUE,
               names_to  = "diagnosis", 
               values_to = "raduis_mean"))
```
I again put it tidy format using pivot_longer.

<!----------------------------------------------------------------------------->

### 2.3 (5 points)

Now, you should be more familiar with your data, and also have made progress in answering your research questions. Based on your interest, and your analyses, pick 2 of the 4 research questions to continue your analysis in milestone 3, and explain your decision. 

Try to choose a version of your data that you think will be appropriate to answer these 2 questions in milestone 3. Use between 4 and 8 functions that we've covered so far (i.e. by filtering, cleaning, tidy'ing, dropping irrelvant columns, etc.). Save this tibble as an `.RDS` object with the function `saveRDS()`. 

<!--------------------------- Start your work below --------------------------->

```{r}
(cancer_sample_two <- cancer_sample[,c(1:12)]  %>% 
   drop_na()%>% 
   filter(radius_mean<22)
 )
saveRDS(cancer_sample_two, file = "mydataframe.rds")
```

For now, I only found 3 appropriate functions to apply to this data frame.
<!----------------------------------------------------------------------------->

*When you are done, knit an `md` file. This is what we will mark! Make sure to open it and check that everything has knitted correctly before submitting your tagged release.*




